Decompositions of congruence subgroups \\ of Chevalley groups

### 2018-02-12 paper.tex
More fixes

Let $J\subseteq \Pi$ be a subset of fundamental roots. Define
\begin{align*}
& \shape(J, \beta)=\sum\nolimits_{i\in J} m_i(\beta) \alpha_i, \\
& \Delta_J = \{\alpha \in \Phi \mid \shape(J, \alpha)=0\}, \\
& \Sigma^\pm_J = \{\alpha \in \Phi \mid \shape(J, \alpha) \in \mathbb{Z}_{\geqslant0} \Phi^\pm \}, \\
& S_J^\pm = \Delta_J \sqcup \Sigma_J^\pm.
\end{align*}
Clearly, $\Delta_J$ is a reductive subset, while $S^\pm_J$ and $\Sigma^\pm_J$ are parabolic and special subsets, respectively.
For two disjoint subsets $I, J\subseteq \Pi$ one has 
\[ \Sigma^\pm_{I \cup J} = \Sigma^\pm_I\cup\Sigma^\pm_J, \quad \Delta_{I\cup J} = \Delta_I \cap \Delta_J. \]
We omit curly braces in the above notations when $J$ is a one- or two-element set, e.\,g. $\Delta_k=\Delta_{\{k\}}$ and $\Sigma_{i, j}^\pm=\Sigma_{\{i, j\}}^\pm$, etc.
We call $\shape(\{s\}, \beta)$ simply the $s$-shape of $\beta$.

@@

\section{Relative parabolic factorizations} \label{sec:factorizations}
In this section we formulate and prove the relative versions of the decompositions from~\cite{St78}, which will be the main technical tools throughout the next section.

\subsection{Relative Dennis--Vaserstein decompositions}\label{sec:dennis-vaserstein}
Let $\Phi$ be an irreducible root system of rank $\ell$.
In what follows, $r$ and $s$ are two distinct integers such that $1 \leqslant r, s \leqslant \ell$.
From the Levi decomposition~\eqref{rel:Levi-decomp} it follows that the following four subsets of $\E(\Phi, R, I)$ are equal:
\begin{multline*}
A_{rs} \coloneqq \U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \E(\Delta_r, R, I) \cdot \EP_s(R, I) = \\
= \U(\Sigma_r^+, I)\cdot \U(\Sigma^-_r, I) \cdot \E(\Delta_r, R, I) \cdot \EP_s(R, I) = \hspace{5em} \\
\hspace{5em} = \EP_r(R, I) \cdot \E(\Delta_s, R, I) \cdot \U(\Sigma_s^-, I)\cdot \U(\Sigma_s^+, I) = \\
= \EP_r(R, I) \cdot \U(\Sigma^-_r \cap \Sigma^-_s, I) \cdot \EP_s(R, I).
\end{multline*}

Denote by $N_{rs}$ the subgroup of $\E(\Phi, R)$, consisting of elements normalizing the above subset:
\[ N_{rs} = \left\{ g\in \E(\Phi, R)\ \middle|\ g \cdot A_{rs} \cdot g^{-1} \cup g^{-1} \cdot A_{rs} \cdot g \subseteq A_{rs}\right\}. \]

\begin{lemma}\label{lemma:dv-normal} 
For every $\alpha \in \Delta_{r, s} \cup \Phi^+ $ one has $X_\alpha(R) \subseteq N_{rs}$. \end{lemma}
\begin{proof}
Notice that for every $i$ the group $\EP_i(R, I)$ is normalized by $\EP_i(R)$, hence it is normalized by $X_\alpha(R)$, $\alpha \in S_i^+$.
Since $\E(\Delta_{r, s}, R)$ normalizes $\U(\Sigma_r^- \cap \Sigma_s^-, I)$, we obtain the assertion for $\alpha \in \Delta_{r, s}$.

Now let $\alpha$ be a positive simple root. It is clear that $\alpha$ lies either in $\Delta_r$ or $\Delta_s$.
Assume, for example, the latter. From the Levi decomposition we get $\U(\Sigma_r^- \cap \Sigma_s^-, I)^{X_\alpha(R)} \subseteq \U(\Sigma_s^-, I) \subseteq A_{rs}$,
consequently $X_{\alpha}(R) \subseteq N_{rs}$.
Now it is clear that $N_{rs}$ contains the subgroup $\U(\Phi^+, R)$ generated by $X_{\alpha}(R)$, $\alpha\in\Pi$, which implies the assertion.
\end{proof}

\begin{proof} [Proof of \cref{thm:DennisVaserstein}]
First of all, notice that Dennis--Vaserstein decomposition follows the stronger Gauss decomposition (see~\cref{thm:srRI1} below).
Thus, we may assume that $\ell>2$ and, in particular, that $\Phi\neq \mathsf{G}_2$.

Since the equality $A_{rs} = \E(\Phi, R, I)$ implies $A_{sr} = \E(\Phi, R, I)$, it suffices to consider only the following possibilities for $\Phi$, $s$, $r$: % listed in~\cref{table:dv-reps}.
\begin{table}[htb]
\[\begin{array}{l @{\qquad} l @{\qquad} c @{\quad} c @{\quad} c @{\qquad} c @{\qquad} c @{\qquad} c}
\Phi                                 & s    &r      & |\Lambda(\pi)| & \text{type of $\pi$} & \text{type of $\Delta_r$} & |\Lambda(\pi')|& |\Gamma|  \\ \hline\vphantom{\Bigl(}
\rA_\ell, \ \ell\geqslant 2           & 1    &\ell   & \ell+1         & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rB_\ell, \ \ell\geqslant 2           & 1    &\ell   & 2\ell+1        & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rC_\ell, \ \ell\geqslant 2           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\
\rD_\ell, \ \ell\geqslant 4           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 2  \\ 
\rD_\ell, \ \ell\geqslant 4           & \ell &\ell-1 & 2^{\ell-1}     & \text{half-spinor}   & \rA_{\ell-1}              & \ell           & \ell-2  \\
\rE_\ell, \ \ell=6, 7, 8              & \ell &2      & 27, 56, 248    & \text{minimal}       & \rA_{\ell-1}              & \ell           & 3       \\ 
\rE_\ell, \ \ell=6, 7, 8              & \ell &1      & 27, 56, 248    & \text{minimal}       & \rD_{\ell-1}              & 2(\ell-1)      & \ell-1 \\
\rF_4,                                 & 4    & 1     & 26             & \text{minimal}       & \rC_{3}                  & 6              & 3 \end{array}\]
 \caption{List of representations used in the proof of \cref{thm:DennisVaserstein}.} \label{table:dv-reps}
\end{table}

Denote by $L_{rs}$ the monoid consisting of elements $g\in \E(\Phi, R, I)$ such that $g \cdot A_{rs} \subseteq A_{rs}$.
It is easy to see that $L_{rs}$ contains $\EP_r(R, I)$ and is normalized by $N_{rs}$, i.\,e. ${L_{rs}}^{N_{rs}}\subseteq L_{rs}$. Indeed, for $g\in N_{rs}$, $h\in L_{rs}$ one has
\begin{equation}\label{rel:NnormL} h^g \cdot A_{rs} = g^{-1} \cdot h \cdot g \cdot A_{rs} \cdot g^{-1} \cdot g \subseteq g^{-1} \cdot h \cdot A_{rs} \cdot g \subseteq {A_{rs}}^g \subseteq A_{rs}.\end{equation}

\cref{lemma:Stein_reduction} below guarantees that $N_{rs}$ contains the root subgroup $X_{-\alpha_{r}}(R)$, which together with~\cref{lemma:dv-normal} 
 implies that $\EP_s(R) \subseteq N_{rs}$ and consequently, that $\EP_s(R, I) \subseteq L_{rs}$. Indeed for $g\in \EP_s(R, I) \subseteq \EP_s(R)$ one has
$g \cdot L_{rs} = g \cdot L_{rs} \cdot g^{-1} \cdot g \subseteq L_{rs} \cdot g \subseteq L_{rs}.$

Our next goal is to verify the inclusion $\U(\Sigma_s^-, I) \subseteq L_{rs}$. Our main tools will be formula~\eqref{rel:NnormL} and \cref{item-trans2}.
Since $\widetilde{W}(\Delta_s) \subseteq \E(\Delta_s, R) \subseteq N_{rs}$ and 
 $X_{\alpha}(I) \subseteq L_{rs}$ for $\alpha \in \Sigma^-_s\cap \Delta_r$, 
we immediately obtain that $L_{rs}$ contains root subgroups $X_\alpha(I)$ corresponding to $\alpha$
 lying in the orbit $O = W(\Delta_s) \cdot (\Sigma^-_s\cap \Delta_r)$.
Thus, we are left to consider four cases in which $O$ does not coincide with $\Sigma^-_s$. 
\begin{enumerate} 
 \item \textit{Case $\Phi = \rE_8$, $s=8$, $r=1,2$.}  
 There is only one root in $\Sigma_8^-$ not lying in $O$, namely $-\alpha_{\max}$ with $\shape(8,-\alpha_{\max}) = -2$.
 Since $X_{-\alpha_8}(R) \subseteq N_{rs}$, $X_{\alpha_8-\alpha_{\max}}(I) \subseteq L_{rs}$ the desired inclusion
  follows from Chevalley commutator formula $x_{\alpha_{\max}}(\pm ab)=[x_{\alpha_8-\alpha_{\max}}(a), x_{\alpha_8}(b)].$

 \item \textit{Case $\Phi = \rB_\ell$, $s=1$, $r=\ell$.} 
 Only the sole short root of $\Sigma_1^-$ is not contained in $O$. Consider the following two roots of $\Phi$:
 \begin{equation}\nonumber \beta = \alpha_2 + \ldots + \alpha_\ell,\ \gamma = -\alpha_1 - 2\alpha_2 - \ldots -2\alpha_\ell.\end{equation}
 We need to verify $X_{\beta+\gamma}(I) \subseteq L_{rs}$, but this follows from $X_{\gamma}(I) \cup X_{-\alpha_1}(I) \subseteq L_{rs}$, $X_{\beta}(R) \subseteq N_{rs}$ using Chevalley commutator formula:
  \begin{equation} \label{eq:CCF} x_{\beta+\gamma}(-ab) = [x_{\gamma}(a),\, x_{\beta}(b)] \cdot x_{\gamma + 2\beta}(ab^2). \end{equation}
 \item \textit{Case $\Phi = \rF_4$, $s=4$, $r=1$.} There are three possibilities for the $s$-shape and length of $\alpha\in \Sigma_4^-$: 
 \begin{itemize}
  \item $\shape(4, \alpha) = 1$, $\alpha$ is short (there are 8 such roots);
  \item $\shape(4, \alpha) = 2$, $\alpha$ is long (there are 6 such roots);
  \item $\shape(4, \alpha) = 2$, $\alpha$ is short (there is only one such root: $\delta = -\alpha_1 -2\alpha_2-3\alpha_3-2\alpha_4$).
 \end{itemize}
 Since $\Delta_r \cap \Sigma_s^-$ contains roots of the first two types, we 
  only need to verify $X_\delta(I) \subseteq L_{rs}$, which again follows from~\eqref{eq:CCF} with $\beta = \alpha_2 + \alpha_3,\ \gamma = -\alpha_{{\max}}$. 
 \item \textit{Case $\Phi = \rC_\ell$, $s=1$, $r=\ell$.} 
  In this case the inclusion $X_{-\alpha_{\max}}(I) \subseteq L_{rs}$ does not immediately follow from commutator formulae
   and we need to invoke stability conditions once again.
  Using~\cref{lemma:DVST} below we obtain that $N_{rs}$ contains $X_{-\alpha_1}(R)$ and therefore coincides with $\E(\Phi, R)$.
  The required inclusion now follows from this and~\cref{item-trans1}.
\end{enumerate}

Thus, we have just shown that $\mathcal{X}\subseteq L_{rs}$. 
Recall from~\cite[Theorem~3.4]{S} that the group $\E(\Phi, R, I)$ is generated by $\mathcal{Z}(\Sigma_s^-)$.
On the other hand, from $\mathcal{Z}(\Sigma_s^-) \subseteq \mathcal{X}^{\EP_s(R)} \subseteq L_{rs}^{N_{rs}} \subseteq L_{rs}$ we conclude that 
 $\E(\Phi, R, I) \subseteq L_{rs}$, which completes the proof.
\end{proof}

\begin{lemma}\label{lemma:dv_unipotent} For any $1\leq i\leq n$ the following statements hold. 
\begin{thmlist} \item \label{item-dvu1} $\U(\Phi^+, I) = X_{\alpha_{i}}(I)\cdot \U(\Phi^+\setminus\{\alpha_{i}\}, I) = \U(\Phi^+\setminus\{\alpha_{i}\}, I)\cdot X_{\alpha_{i}}(I)$;
\item \label{item-dvu2} For any $\xi\in R$ one has $\U(\Phi^+\setminus\{\alpha_i\}, I)^{x_{-\alpha_{i}}(\xi)} \subseteq \U(\Phi^+, I)$;
\item \label{item-dvu3} $\U(\Phi^+, I)\cdot \U(\Phi^-, I) \subseteq \U(\Phi^+\setminus \{\alpha_i\}, I) \cdot \U(\Phi^-, I) \cdot X_{\alpha_{i}}(I) \cdot X_{-\alpha_{i}}(I)$.
\end{thmlist} \end{lemma}
\begin{proof} The first two statements easily follow from the Chevalley commutator formula, while the third one is a formal consequence of the first two. \end{proof}

@@

\subsection{Relative Gauss decomposition}\label{sec:gauss}
The proof of the Gauss decomposition presented below is similar to the proof in the absolute case (cf.~\cite[Theorem~5.1]{Sm12}).
\begin{prop}\label{thm:Gauss}
Let $\Phi$ be a reduced irreducible root system of rank $\ell$ and let $\Delta_1$, $\Delta_\ell$ be
two reductive subsystems of $\Phi$, corresponding to the endnodes of the Dynkin diagram of $\Phi$.
Suppose that both relative elementary subgroups $\E(\Delta_i, R, I)$, $i=1, \ell$ admit triangular decomposition with $N$ triangular factors:
\[ \E(\Delta_i, R, I) = \Hh(\Delta_i, R, I) \cdot \U(\Delta^+_i, I) \cdot \U(\Delta^-_i, I) \cdot \ldots \cdot \U(\Delta^\pm_i, I), \quad i=1, \ell. \]
Then the group $\E(\Phi, R, I)$ admits decomposition with the same number of factors:
\[ \E(\Phi, R, I) = \Hh(\Phi, R, I) \cdot \U(\Phi^+, I) \cdot \U(\Phi^-, I) \cdot \ldots \cdot \U(\Phi^\pm, I). \]
\end{prop}
\begin{proof}
Denote by $Y$ the product of subgroups in the right-hand side of the above equality.
To show that $Y=\E(\Phi, R, I)$ we need to verify the following assertions:
\begin{enumerate}
\item $Y$ is normalized by $\E(\Phi, R)$, i.\,e. $Y^{\E(\Phi, R)}\subseteq Y$;
\item there exists a set $X$, generating $\E(\Phi, R, I)$ as a \emph{normal} subgroup of $\E(\Phi, R)$ such that $XY\subseteq Y$.
\end{enumerate}
To prove the first assertion it suffices to show that $Y^{x_\alpha(\xi)} \subseteq Y$ for any $\alpha\in \pm \Pi$, $\xi\in R$.
Fix a root $\alpha\in\pm\Pi$. Clearly, $\alpha \in \Delta_i$ for $i=1$ or for $i=\ell$, and we can factor $Y$ as follows:
\[ Y=\Hh(\Phi, R, I) \cdot \U(\Delta_i^+, I) \cdot \ldots \cdot \U^\pm(\Delta_i, I) \cdot \U(\Sigma_i^+, I) \cdot \ldots \cdot \U(\Sigma_i^\pm, I). \]
For every  $h \in \Hh(\Phi, R, I)$ for some $s\in I$ one has
\[ x_\alpha(\xi)\cdot h = h\cdot x_\alpha((1+s)\xi)=h\cdot x_\alpha(\xi)\, x_\alpha(s\xi). \]

>>>>

Let $J\subseteq \Pi$ be a subset of fundamental roots. Define
\begin{align*}
& \shape(J, \beta)=\sum\nolimits_{i\in J} m_i(\beta) \alpha_i, \\
& \Delta_J = \{\alpha \in \Phi \mid \shape(J, \alpha)=0\}, \\
& \Sigma^\pm_J = \{\alpha \in \Phi \mid \shape(J, \alpha) \in \mathbb{Z}_{\geqslant0} \Phi^\pm \}, \\
& S_J^\pm = \Delta_J \sqcup \Sigma_J^\pm.
\end{align*}
Clearly, $\Delta_J$ is a reductive subset, while $S^\pm_J$ and $\Sigma^\pm_J$ are parabolic and special subsets, respectively.
For two disjoint subsets $I, J\subseteq \Pi$ one has 
\[ \Sigma^\pm_{I \cup J} = \Sigma^\pm_I\cup\Sigma^\pm_J, \quad \Delta_{I\cup J} = \Delta_I \cap \Delta_J. \]
We omit curly braces in the above notations when $J$ is a one- or two-element set, e.\,g. $\Delta_k=\Delta_{\{k\}}$ and $\Sigma_{i, j}^\pm=\Sigma_{\{i, j\}}^\pm$, etc.
We call $\shape(\{\alpha_s\}, \beta)$ simply the $s$-shape of $\beta$.

@@

\section{Relative parabolic factorizations} \label{sec:factorizations}
In this section we formulate and prove relative versions of the decompositions from~\cite{St78}, which will be our main technical tools throughout the next section.

Let $G$ be a group and $A$ its subset. Denote by $L(A)$, $R(A)$ and $N(A)$ the left and the right stabilizers of $A$ and the normalizer of $A$:
\begin{align*} 
 L(A) & =  \left\{ g\in G\ \middle |\ g \cdot A \cup g^{-1} \cdot A \subseteq A \right\}\!,\\
 R(A) & =  \left\{ g\in G\ \middle |\ A \cdot g \cup A \cdot g^{-1} \subseteq A \right\}\!,\\
 N(A) & =  \left\{ g\in G\ \middle |\ g \cdot A \cdot g^{-1} \cup g^{-1} \cdot A \cdot g \subseteq A \right\}.
\end{align*}
It is easy to see that $L(A)$, $R(A)$ and $N(A)$ are subgroups of $G$. The following obvious lemma will be useful in the sequel.
\begin{lemma} \label{lem:LN} The subgroup $L(A)$ is normalized by $N(A)$. Moreover, one has $R(A) \cap N(A) \subseteq L(A)$, $L(A) \cap N(A) \subseteq R(A)$. \end{lemma}
\begin{comment}
\begin{proof}
Let $h \in L(A)$, $g \in N(A)$. The first claim follows from the following chain of inclusions:
\begin{equation} h^g \cdot A = g^{-1} \cdot h \cdot g \cdot A \cdot g^{-1} \cdot g \subseteq g^{-1} \cdot h \cdot A \cdot g \subseteq {A}^g \subseteq A. \end{equation} 
Now if, in addition, $g \in R(A)$ we obtain
$g \cdot A = g \cdot A \cdot g^{-1} \cdot g \subseteq A \cdot g \subseteq A,$ which implies the second claim.
\end{proof}
\end{comment}

\subsection{Relative Dennis--Vaserstein decompositions}\label{sec:dennis-vaserstein}
Let $\Phi$ be an irreducible root system of rank $\ell$ and $r$ and $s$ be two distinct integers such that $1 \leqslant r, s \leqslant \ell$.
From Levi decomposition~\eqref{rel:Levi-decomp} it follows that the following four subsets of $\E(\Phi, R, I)$ are equal:
\begin{multline*}
\U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \E(\Delta_r, R, I) \cdot \EP_s(R, I) = \\
= \U(\Sigma_r^+, I)\cdot \U(\Sigma^-_r, I) \cdot \E(\Delta_r, R, I) \cdot \EP_s(R, I) = \hspace{5em} \\
\hspace{5em} = \EP_r(R, I) \cdot \E(\Delta_s, R, I) \cdot \U(\Sigma_s^-, I)\cdot \U(\Sigma_s^+, I) = \\
= \EP_r(R, I) \cdot \U(\Sigma^-_r \cap \Sigma^-_s, I) \cdot \EP_s(R, I).
\end{multline*}
Denote the above subset by $A_{rs}$.

\begin{lemma}\label{lemma:dv-normal} 
For every $\alpha \in \Delta_{r, s} \cup \Phi^+ $ one has $X_\alpha(R) \subseteq N(A_{rs})$. \end{lemma}
\begin{proof}
Notice that for every $i$ the group $\EP_i(R, I)$ is normalized by $\EP_i(R)$, hence it is normalized by $X_\alpha(R)$, $\alpha \in S_i^+$.
Since $\E(\Delta_{r, s}, R)$ normalizes $\U(\Sigma_r^- \cap \Sigma_s^-, I)$, we obtain the required assertion for $\alpha \in \Delta_{r, s}$.

Now let $\alpha$ be a positive simple root. It is clear that $\alpha$ lies either in $\Delta_r$ or $\Delta_s$.
Assume, for example, the latter. From the Levi decomposition we obtain $\U(\Sigma_r^- \cap \Sigma_s^-, I)^{X_\alpha(R)} \subseteq \U(\Sigma_s^-, I) \subseteq A_{rs}$, therefore $X_{\alpha}(R) \subseteq N(A_{rs})$.
Thus, $N(A_{rs})$ contains the subgroup $\U(\Phi^+, R)$ generated by $X_{\alpha}(R)$, $\alpha\in\Pi$, which completes the proof.
\end{proof}

\begin{proof} [Proof of \cref{thm:DennisVaserstein}]
First of all, notice that in the case $\ell=2$, $\sr(I)=1$ Dennis--Vaserstein decomposition can be deduced from Gauss decomposition given by~\cref{thm:srRI1} below.
Thus, without loss of generality we may assume $\ell>2$ and, in particular, that $\Phi\neq \mathsf{G}_2$.

Since the equality $A_{rs} = \E(\Phi, R, I)$ implies $A_{sr} = \E(\Phi, R, I)$, it suffices to consider only the possibilities for $\Phi$, $s$, $r$ listed in~\cref{table:dv-reps} below.
\begin{table}[htb]
\[\begin{array}{l @{\qquad} l @{\qquad} c @{\quad} c @{\quad} c @{\qquad} c @{\qquad} c @{\qquad} c}
\Phi                                 & s    &r      & |\Lambda(\pi)| & \text{type of $\pi$} & \text{type of $\Delta_r$} & |\Lambda(\pi')|& |\Gamma|  \\ \hline\vphantom{\Bigl(}
\rA_\ell, \ \ell\geqslant 2           & 1    &\ell   & \ell+1         & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rB_\ell, \ \ell\geqslant 2           & 1    &\ell   & 2\ell+1        & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\     
\rC_\ell, \ \ell\geqslant 2           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 1  \\
\rD_\ell, \ \ell\geqslant 4           & 1    &\ell   & 2\ell          & \text{natural}       & \rA_{\ell-1}              & \ell           & 2  \\ 
\rD_\ell, \ \ell\geqslant 4           & \ell &\ell-1 & 2^{\ell-1}     & \text{half-spinor}   & \rA_{\ell-1}              & \ell           & \ell-2  \\
\rE_\ell, \ \ell=6, 7, 8              & \ell &2      & 27, 56, 248    & \text{minimal}       & \rA_{\ell-1}              & \ell           & 3       \\ 
\rE_\ell, \ \ell=6, 7, 8              & \ell &1      & 27, 56, 248    & \text{minimal}       & \rD_{\ell-1}              & 2(\ell-1)      & \ell-1 \\
\rE_\ell, \ \ell=6, 7, 8              & 1    &2      & 27, 133, 3875  & \text{???}           & \rA_{\ell-1}              & ?              & ? \\
\rF_4,                                & 4    & 1     & 26             & \text{minimal}       & \rC_{3}                   & 6              & 3 \end{array}\]
 \caption{List of representations used in the proof of \cref{thm:DennisVaserstein}.} \label{table:dv-reps}
\end{table}

From \cref{lemma:Stein_reduction} below it follows that $N(A_{rs})$ contains root subgroup $X_{-\alpha_{r}}(R)$, which together with~\cref{lemma:dv-normal} 
 implies that $\EP_s(R) \subseteq N(A_{rs})$, and hence by~\cref{lem:LN} that $\EP_s(R, I) \subseteq L(A_{rs})$.

Our next goal is to verify the inclusion $\U(\Sigma_s^-, I) \subseteq L(A_{rs})$. Our main tools will be~\cref{lem:LN} and \cref{item-trans2}.
Since $\widetilde{W}(\Delta_s) \subseteq \E(\Delta_s, R) \subseteq N(A_{rs})$ and 
 $X_{\alpha}(I) \subseteq L(A_{rs})$ for $\alpha \in \Sigma^-_s\cap \Delta_r$, 
we immediately obtain that $L(A_{rs})$ contains root subgroups $X_\alpha(I)$ for all $\alpha$ lying in the orbit $O = W(\Delta_s) \cdot (\Sigma^-_s\cap \Delta_r)$.
Thus, we are left to consider four cases in which $O$ does not coincide with $\Sigma^-_s$.
\begin{enumerate} 
 \item \textit{Case $\Phi = \rE_8$, $s=8$, $r=1,2$.}  
 There is only one root in $\Sigma_8^-$ not lying in $O$, namely $-\alpha_{\max}$ with $\shape(\{\alpha_8\}, -\alpha_{\max}) = -2$.
 Since $X_{-\alpha_8}(R) \subseteq N(A_{rs})$, $X_{\alpha_8-\alpha_{\max}}(I) \subseteq L(A_{rs})$ the desired inclusion
  follows from Chevalley commutator formula $x_{\alpha_{\max}}(\pm ab)=[x_{\alpha_8-\alpha_{\max}}(a), x_{\alpha_8}(b)].$
 \item \textit{Case $\Phi = \rE_7, \rE_8$, $s=1$, $r=2$.} ???
 \item \textit{Case $\Phi = \rB_\ell$, $s=1$, $r=\ell$.} 
 Only the sole short root of $\Sigma_1^-$ is not contained in $O$. Consider the following two roots of $\Phi$:
 \begin{equation}\nonumber \beta = \alpha_2 + \ldots + \alpha_\ell,\ \gamma = -\alpha_1 - 2\alpha_2 - \ldots -2\alpha_\ell.\end{equation}
 We need to verify $X_{\beta+\gamma}(I) \subseteq L(A_{rs})$, but this follows from $X_{\gamma}(I) \cup X_{-\alpha_1}(I) \subseteq L_{rs}$,
  $X_{\beta}(R) \subseteq N(A_{rs})$ using Chevalley commutator formula:
  \begin{equation} \label{eq:CCF} x_{\beta+\gamma}(-ab) = [x_{\gamma}(a),\, x_{\beta}(b)] \cdot x_{\gamma + 2\beta}(ab^2). \end{equation}
 \item \textit{Case $\Phi = \rF_4$, $s=4$, $r=1$.} There are three possibilities for the $s$-shape and length of $\alpha\in \Sigma_4^-$: 
 \begin{itemize}
  \item $\shape(\{\alpha_4\}, \alpha) = 1$, $\alpha$ is short (there are 8 such roots);
  \item $\shape(\{\alpha_4\}, \alpha) = 2$, $\alpha$ is long (there are 6 such roots);
  \item $\shape(\{\alpha_4\}, \alpha) = 2$, $\alpha$ is short (there is only one such root: $\delta = -\alpha_1 -2\alpha_2-3\alpha_3-2\alpha_4$).
 \end{itemize}
 Since $\Delta_r \cap \Sigma_s^-$ contains roots of the first two types, we only need to verify $X_\delta(I) \subseteq L(A_{rs})$. 
 But the latter inclusion follows from formula~\eqref{eq:CCF} with $\beta = \alpha_2 + \alpha_3,\ \gamma = -\alpha_{{\max}}$. 
 \item \textit{Case $\Phi = \rC_\ell$, $s=1$, $r=\ell$.} 
  In this case the inclusion $X_{-\alpha_{\max}}(I) \subseteq L(A_{rs})$ does not follow from commutator formulae, therefore we need to invoke stability assumptions one more time.
  From~\cref{lemma:DVST} below we obtain that $N(A_{rs})$ contains $X_{-\alpha_1}(R)$ and therefore coincides with $\E(\Phi, R)$.
  The required inclusion now follows~\cref{item-trans1}.
\end{enumerate}

Thus, we have just shown that $\mathcal{X}\subseteq L(A_{rs})$. 
Recall from~\cite[Theorem~3.4]{S} that the group $\E(\Phi, R, I)$ is generated by the subset $\mathcal{Z}(\Sigma_s^-)$.
On the other hand, from $\mathcal{Z}(\Sigma_s^-) \subseteq \mathcal{X}^{\EP_s(R)} \subseteq L(A_{rs})^{N(A_{rs})} \subseteq L(A_{rs})$ we conclude that 
 $\E(\Phi, R, I) \subseteq L(A_{rs})$, which completes the proof.
\end{proof}

\begin{lemma}\label{lemma:dv_unipotent} For any $1\leq i\leq n$ the following statements hold. 
\begin{thmlist} \item \label{item-dvu1} $\U(\Phi^+, I) = X_{\alpha_{i}}(I)\cdot \U(\Phi^+\setminus\{\alpha_{i}\}, I) = \U(\Phi^+\setminus\{\alpha_{i}\}, I)\cdot X_{\alpha_{i}}(I)$;
\item \label{item-dvu2} For any $\xi\in R$ one has $\U(\Phi^+\setminus\{\alpha_i\}, I)^{x_{-\alpha_{i}}(\xi)} \subseteq \U(\Phi^+, I)$;
\item \label{item-dvu3} $\U(\Phi^+, I)\cdot \U(\Phi^-, I) \subseteq \U(\Phi^+\setminus \{\alpha_i\}, I) \cdot \U(\Phi^-, I) \cdot X_{\alpha_{i}}(I) \cdot X_{-\alpha_{i}}(I)$.
\end{thmlist} \end{lemma}
\begin{proof} The first two assertions follow from Chevalley commutator formula, while the third one is a consequence of the first two. \end{proof}

@@

\subsection{Relative Gauss decomposition}\label{sec:gauss}
The proof of the Gauss decomposition presented below is similar to the proof in the absolute case (cf.~\cite[Theorem~5.1]{Sm12}).

\begin{prop} \label{thm:Gauss}
Let $\Phi$ be a reduced irreducible root system of rank $\ell > 1$ and let $\Delta_k$, $\Delta_l$ be
the reductive subsystems of $\Phi$ corresponding to a pair of distinct simple roots $\alpha_k$, $\alpha_l \in \Pi$.
Suppose that both relative elementary subgroups $\E(\Delta_i, R, I)$, $i=k, l$ admit Gauss decomposition:
\[ \E(\Delta_i, R, I) = \Hh(\Delta_i, R, I) \cdot \U(\Delta^+_i, I) \cdot \U(\Delta^-_i, I) \cdot \U(\Delta^+_i, I), \quad i=k, l. \]
Then $\E(\Phi, R, I)$ also admits Gauss decomposition:
\begin{equation} \label{eq:gauss1} \E(\Phi, R, I) = \Hh(\Phi, R, I) \cdot \U(\Phi^+, I) \cdot \U(\Phi^-, I) \cdot \U(\Phi^+, I). \end{equation}
\end{prop}
\begin{proof}
For a closed subset of roots $S \subseteq \Phi^+$ set $A(S, I) = \U(S, I) \cdot \U(-S, I) \cdot \U(S, I)$ (here by $-S$ we denote the corresponding subset of opposite roots).
Notice that from Levi decomposition it follows that $A(\Phi^+, I) = A(\Delta_i^+, I) \cdot A(\Sigma_i^+, I)$.
